# Hypothesis

```{r,include=FALSE}
library(knitr)
opts_chunk$set(tidy = TRUE, tidy.opts = list(width.cutoff = 40))
```

## Density Plot for Height and Weight

```{r, include=FALSE}
library(crayon)
library(ggplot2)
library(gridExtra)
library(reshape2)
```

### Density Plot for Height Distribution Over Seasons

```{r, fig.width=11, fig.height=5}
heightplot <- ggplot(nba, aes(x = player_height)) +
  geom_density(aes(fill = season), alpha = 0.4) +
  geom_vline(aes(xintercept = mean(player_height)), linetype = "dashed", color = "red") +
  ggtitle("Density Plot of Player Heights") +
  xlab("Player Height (cm)") +
  ylab("Density") +
  theme_minimal()
print(heightplot)
```

#### Summary

- A more noticeable shift in the mean height of players over seasons. The mean height has decreased steadily.

- A narrower range of heights among players in the recent NBA compared to the past.

- There's a noticeable peak around the 200-210 cm range, suggesting that a significant number of players fall within this height bracket.

### Density Plot for Weight Distribution Over Seasons

```{r, fig.width=11, fig.height=5}
weightplot <- ggplot(nba, aes(x = player_weight)) +
  geom_density(aes(fill = season), alpha = 0.4) +
  geom_vline(aes(xintercept = mean(player_weight)), linetype = "dashed", color = "red") +
  ggtitle("Density Plot of Player Weights") +
  xlab("Player Weight (kg)") +
  ylab("Density") +
  theme_minimal()
print(weightplot)
```

#### Summary

- The mean weight seems to have shifted towards the left, suggesting that players have, on average, become lighter over the years.
- The distributions for the recent seasons appear narrower, indicating less variability in player weights now than in earlier seasons.

### General Conclusion

In summary, over the years, NBA players have, on average, become shorter and lighter, with a narrower range of both weights and heights represented in recent seasons.

## Inference & Hypothesis Based On The Density Plot and Initial Summary

- **H1** Versatility & Position-less Basketball: the narrower range of heights and weights indicates that there might be an increasing trend of “position-less” basketball, specifically players are no longer strictly confined to traditional roles based on their physical attributes.

- **H2** Evolution in Playing Style: the traditional center-focused style of play cannot adapt to the pace of the modern NBA.