# Correlation Analysis

```{r, include=FALSE}
selected_vars <- c('player_height', 'player_weight', 'reb', 'ast', 'net_rating', 'oreb_pct', 'dreb_pct', 'usg_pct', 'ts_pct', 'ast_pct')
nba_selected <- subset(nba, select = selected_vars)
cor_matrix <- cor(nba_selected, use = "complete.obs")
melted_cormat <- melt(cor_matrix)
```

## Correlation Between H/W and Others

- The correlation matrix visually represents the relationship between height/weight and other basketball metrics. 
- Correlations range between -1.0 and 1.0, with darker red indicating a strong positive correlation, blue indicating a strong negative correlation, and lighter shades (towards white) indicating weaker or no correlation.

```{r, fig.width=11, fig.height=5}
ggplot(data = melted_cormat, aes(x=Var1, y=Var2)) +
  geom_tile(aes(fill=value), color='white') +
  scale_fill_gradient2(low='blue', high='red', mid='white', midpoint=0, limit=c(-1,1), space='Lab', name='Correlation') +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, vjust=1, size=12, hjust=1),
        axis.text.y=element_text(size=12)) +
  coord_fixed()
```

## Correlation Matrix

- A strong positive correlation between height and defensive rebound percentage as well as offensive rebound percentage.
- Weight also shows positive correlations with both rebound percentage, though slightly weaker compared to height.
- Height seems to have a negative correlation with assist percentage, suggesting taller players might assist less.
- Other metrics like true shooting percentage and usage percentage have weaker correlations with height and weight.

## Correlation Between H/W and R/A

Four scatter plots depict the relationships between player height, player weight, rebounds, and assists.

```{r}
height_reb<-ggplot(nba,aes(x=player_height,y=reb))
height_reb<-height_reb+geom_point()

height_ast<-ggplot(nba,aes(x=player_height,y=ast))
height_ast<-height_ast+geom_point()

weight_reb<-ggplot(nba,aes(x=player_weight,y=reb))
weight_reb<-weight_reb+geom_point()

weight_ast<-ggplot(nba,aes(x=player_weight,y=ast))
weight_ast<-weight_ast+geom_point()
```

```{r, fig.width=11, fig.height=5}
grid.arrange(height_reb+geom_smooth(), height_ast+geom_smooth(), weight_reb+geom_smooth(), weight_ast+geom_smooth(), ncol = 2)
```

```{r,include=FALSE}
print("hello world")
```
```{r,include=FALSE}
print("hello world")
```
```{r,include=FALSE}
print("hello world")
```

## Summary From Scatter Plots

- Consistent with the correlation matrix, height and rebound idsplays an upward trend, indicating that taller players tend to have more rebounds.
- Conversely, height and assist depicts a downward trend, signifying that shorter players usually have more assists.

## Net Rating By Player's Physical Attributes

The boxplots depict the distribution of net rating (i.e., the team's point differential per 100 possessions while a player is on court) across various player heights and weights.

```{r, fig.width=11, fig.height=5}
heightdatapart <- data.frame(player_height = nba$player_height, net_rating = nba$net_rating)
heightdatapart <- heightdatapart[(heightdatapart$net_rating >= -50 & heightdatapart$net_rating <= 50), ]
heightdatapart$group <- cut(heightdatapart$player_height, breaks = 7)
ph <- ggplot(data = heightdatapart, aes(x = group, y = net_rating, fill = group)) +
  geom_boxplot() +
  ggtitle("Net Rating by Player Height") +
  theme_minimal()
print(ph)
weightdatapart <- data.frame(player_weight = nba$player_weight, net_rating = nba$net_rating)
weightdatapart <- weightdatapart[(weightdatapart$net_rating >= -50 & weightdatapart$net_rating <= 50), ]
weightdatapart$group <- cut(weightdatapart$player_weight, breaks = 7)
pw <- ggplot(data = weightdatapart, aes(x = group, y = net_rating, fill = group)) +
  geom_boxplot() +
  ggtitle("Net Rating by Player Weight") +
  theme_minimal()
print(pw)
```

## Summary From Box Plot

### Net Rating by Player Height

- Players in the height range of 190 cm to 203 cm tend to have a wider variation in net rating, with some players in this range even having exceptionally high net ratings.
- Taller players, especially those between 206 cm and 211 cm, have a slightly lower median net rating.
- The shorter players, particularly those below 188 cm, also show a lower median net rating, but with less variability than the taller players.
- Players who are 213 cm and above exhibit a narrower interquartile range (middle 50% of data) but with a few outliers having lower net ratings.

### Net Rating by Player Weight

- Players weighing between 70-85 kg and 100-115 kg display a wider range in net rating, with some outliers, particularly on the higher end.
- Players in the weight range of 85-100 kg have a median net rating that's slightly higher, with a more compact interquartile range, indicating less variability in their performance.
- Heavier players, weighing 119 kg and above, exhibit a narrower interquartile range with a few lower outliers.
From the boxplots, it can be concluded that:

### General Summary
- Players of medium height (190 cm to 203 cm) tend to have varied performances, with some players being exceptionally good based on net rating.
- The shortest and the tallest players tend to have a slightly lower median net rating.
- Players of medium weight (85-100 kg) generally have consistent performances, while players on the lighter and heavier ends show greater variability in net rating.
