---
title: "Visualizing NBA"
author: "Junqi Fu"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
output: 
  bookdown::pdf_book:
    latex_engine: xelatex
    keep_tex: true
    number_sections: true
    toc: true
    toc_depth: 3
    fig_caption: true
    dev: cairo_pdf
    latex_options: "scale_down"
    extra_dependencies: subfig

bibliography: [book.bib,packages.bib]
link-citations: yes
colorlinks: yes
github-repo: rstudio/bookdown-demo
---

```{r,include=FALSE}
library(knitr)
opts_chunk$set(tidy = TRUE, tidy.opts = list(width.cutoff = 40))
```

# NBA Data

The data set contains all the players' performance data from season 1996-97 to season 2022-23.

## Data Type

Here's a brief explanation of some important variable in the analysis:

- **player_height**: player's height in the given season.
- **player_weight**: player's weight in the given season.
- **gp**: total games a player has played in the given season.
- **pts**: player's average points per game.
- **reb**: player's average rebound per game.
- **ast**: player's average assist per game.
- **net_rating**: the team's point differential per 100 possessions while a player is on court.
- **oreb_pct**: offensive rebound percentage - an estimate of the percentage of available offensive rebounds a player grabbed.
- **dreb_pct**: defensive rebound percentage - an estimate of the percentage of available defensive rebounds a player grabbed.
- **usg_pct**: usage percentage - an estimate of the percentage of team plays used by a player.
- **ts_pct**: true shooting percentage - a measure of shooting efficiency that takes into account field goals, 3-point field goals, and free throws.
- **ast_pct**: assist percentage - an estimate of the percentage of teammate field goals a player assisted.
- **season**: season - the NBA season for which these stats apply.

## Preliminary Cleaning

```{r,include=FALSE, fig.width=6, fig.height=4}
nba<-read.csv("all_seasons.csv",header=TRUE,sep=",")
sum(is.na(nba))
library(mice)
library(VIM)
```

### Initial Missing Data Imputation

```{r, fig.width=12, fig.height=5}
sum(is.na(nba))
md.pattern(nba)
md.pattern(subset(nba,select=c(player_height,player_weight,gp,pts,reb,ast,net_rating)))
aggr(subset(nba,select=c(player_height,player_weight,gp,pts,reb,ast,net_rating)),prop=F,numbers=TRUE)
```

### Data Cleaning

identify outlier exists in the net_rating, with the 300 as the max and -250 as the min.

```{r, fig.width=11, fig.height=5}
summary(nba)
nba_clean<-(which(nba$net_rating>=300 |nba$net_rating<=-200))
nba<-nba[-nba_clean] #remove the outliers.
```

## Preliminary Descriptive Statistics


-   **age**: the range is between a minimum of 18 years and a maximum of 44 years, with the average age being 26 years old.

```{r, fig.width=11, fig.height=5}
summary(nba$age)
```

-   **height**: the range is between a minimum of 160 cm and a maximum of 231.1 cm, with the average height as 200.7 cm.

```{r, fig.width=11, fig.height=5}
summary(nba$player_height)
```

-   **weight**: the range is between a minimum of 60.33 kg and a maximum of 163.29 kg, with the average as 100 kg.

```{r, fig.width=11, fig.height=5}
summary(nba$player_weight)
```

-   **player**: with a highly competitive threshold, only 2551 players have played in the league since 1996.

```{r, fig.width=11, fig.height=5}
summary(unique(nba$player_name))
```
